Exploring the Effect of Divide Recombine
========================================================
Team: Daniel Bero (dbero), Jennifer Chang (jenchang), Guillermo Basulto-Elias (basulto) 

Exercise 1
--------
Read Andee Kaplan's blog on comparing dplyr and datadr

Run the head to head comparison done by Andee. Do you get similar results?  Using proc.time() might help to do simple timing.

* http://andeekaplan.com/2014/03/09/datadr/

We got the same conclusion: dplyr is faster..
```{r}
library(devtools)
install_github('datadr','hafen')
library(datadr)
data(adult)

## turn adult into a ddf
(adultDdf<-ddf(adult, update=TRUE))
names(adultDdf)
library(ggplot2)
edTable<-summary(adultDdf)$education$freqTable
edTable$var

## Names
names(adultDdf)

## Plot
library(ggplot2)
edTable <- summary(adultDdf)$education$freqTable
edTable$var <- with(edTable, reorder(var, Freq, mean))
qplot(var, Freq, data = edTable, geom = "point") + coord_flip()


datadr_test <- function() {
    # make a preTransFn to group some education levels
    edGroups <- function(v) {
        v$edGroup <- as.character(v$education)
        v$edGroup[v$edGroup %in% c("1st-4th", "5th-6th")] <- "Some-elementary"
        v$edGroup[v$edGroup %in% c("7th-8th", "9th")] <- "Some-middle"
        v$edGroup[v$edGroup %in% c("10th", "11th", "12th")] <- "Some-HS"
        v
    }
    
    # divide by edGroup and filter out 'Preschool'
    byEdGroup <- divide(adultDdf, by = "edGroup", preTransFn = edGroups, filterFn = function(x) x$edGroup[1] != 
        "Preschool", update = TRUE)
    
    print(byEdGroup)
    
    # tabulate number of people in each education group
    edGroupTable <- recombine(byEdGroup, apply = nrow, combine = combRbind())
    print(edGroupTable)
    
    # order for plotting
    edGroupTable$edGroup <- with(edGroupTable, reorder(edGroup, val, mean))
    print(qplot(edGroup, val, data = edGroupTable, geom = "point") + coord_flip())
}
datadr_test()

### The same with plyr
library(dplyr)
dplyr_test <- function() {
    byEdGroup_dplyr <- adult %.% mutate(edGroup = ifelse(as.character(education) %in% 
        c("1st-4th", "5th-6th"), "Some-elementary", ifelse(as.character(education) %in% 
        c("7th-8th", "9th"), "Some-middle", ifelse(as.character(education) %in% 
        c("10th", "11th", "12th"), "Some-HS", as.character(education))))) %.% 
        filter(edGroup != "Preschool") %.% group_by(edGroup)
    print(byEdGroup_dplyr)
    
    edGroupTable_dplyr <- byEdGroup_dplyr %.% summarise(count = n())
    print(edGroupTable_dplyr)
    
    # order for plotting
    edGroupTable_dplyr$edGroup <- with(edGroupTable_dplyr, reorder(edGroup, 
        count, mean))
    print(qplot(edGroup, count, data = edGroupTable_dplyr, geom = "point") + 
        coord_flip())
}
dplyr_test()

library(profr)
system.time({
datadr_prof <- profr(datadr_test())
})

system.time({
dplyr_prof <- profr(dplyr_test())
})
```

Exercise 2
--------
Download the vehicle data from [http://www.fueleconomy.gov/feg/ws/index.shtml]

Your goal here is to compare fitting a glm of city mpg on highway mpg, cylinders, disp, fuelType and year using the full data, against drGLM() and drBLB().  Examine the time taken to compute, and the values of the model estimates.

It will be best to do some cleaning of the data before trying to fit the model, or making the ddf.
* Remove cars listed as year = 2015
* Remove missings on cylinders, displ and make into numeric variables
* Change fuelType into categories "Regular", "Premium", "Diesel", "Electricity", "E85"

```{r}
cars<-read.csv("vehicles.csv", header=TRUE)
cars.a <- subset(cars, year<2015)
cars.b <- subset(cars.a, cylinders!='' & cylinders!='-') # remove records wihtout cylinders data
cars.c <- subset(cars.b, displ!='' & displ!='-')         # remove records without displ data
cars.c$cylinders=as.numeric(as.character(cars.c$cylinders)) # convert cylinders to numeric
cars.c$displ=as.numeric(as.character(cars.c$displ))     # convert displ to numeric

library(plyr)
gastype = cars.c$fuelType
cars.c$gastype = mapvalues(gastype, from=c("CNG", "Gasoline or natural gas", "Gasoline or propane",
                          "Regular", "Midgrade", "Premium", "Diesel", "Electricity",
                          "Premium Gas or Electricity", "Regular Gas and Electricity",
                          "Gasoline or E85", "Premium or E85"), to = c("Regular", 
                          "Regular", "Regular", "Regular", "Regular", "Premium", "Diesel",
                          "Electricity", "Electricity", "Electricity", "E85", "E85"))

cars.d = ddf(cars.c)
time.glm = system.time(glm(UCity ~ UHighway + cylinders + displ + gastype + year, data = cars.c))
time.drGLM = system.time(drGLM(UCity ~ UHighway + cylinders + displ + gastype + year, data = cars.c))
time.drBLB = system.time({cars.d = ddf(cars.c, update=T)
cars.e <- divide(cars.d, by=rrDiv(17000), update=TRUE,postTrans=function(x) 
      x[,c("UCity", "UHighway", "cylinders", "displ", "gastype", "year")])

recombine(cars.e,
   apply=drBLB(
      statistic = function(x, weights)
         coef(glm(UCity ~ UHighway + cylinders + displ + gastype + year, 
            data=x, weights = weights)),
      metric = function(x)
         quantile(x, c(0.05, 0.95)),
      R = 100,
      n = nrow(cars.e)
   ),
   combine=combMean()
)

recombine(
    data = cars.e,
    apply = drGLM(UCity ~ UHighway+cylinders+displ+gastype+year),
    combine = combMeanCoef()
)})

time.glm
time.drGLM
time.drBLB
```

Exercise 3
--------
Modify the MapReduce examples (at http://hafen.github.io/datadr/#mapreduce-examples) to:

Find the make of vehicle that has the lowest mpg

```{r}
minMap <- expression({
   for(curMapVal in map.values)
      collect("min", min(curMapVal$comb08))
      collect("make", curMapVal[which.min(curMapVal$comb08), "make"])
})


minReduce <- expression(
   pre={
      globalMin <- NULL
   },
   reduce={
      globalMin <- unlist(reduce.values)
   },
   post={
      collect(reduce.key, globalMin)
   }
)

minRes <- mrExec(cars.d,
   map = minMap,
   reduce = minReduce
)

minRes[["make"]]
minRes[["min"]]
```
Exercise 4
---------
Write a paragraph on:
* How data cleaning prior to constructing the distributed data objects might be important.
* And how the data setup might affect the types of operations that can be done efficiently


Cleaning data before working with it in any case is important, but especially when working datadr format which is very specific. The ddo and ddf objects are the only objects that many of the datadr functions work well with. If your data isn't clean, it won't be workable. 

<!-- Anything after this is commented out ==================
/```{r}
# Exercise 1 Andee Kaplan example
#library(devtools)
#install_github('datadr','hafen')
#library(datadr)
#data(adult)

# turn adult into a ddf
#(adultDdf<-ddf(adult, update=TRUE))
#names(adultDdf)
#library(ggplot2)
#edTable<-summary(adultDdf)$education$freqTable
#edTable$var

# Exercise 2 Cars
cars<-read.csv("vehicles.csv", header=TRUE)
cars.a <- subset(cars, year<2015)
cars.b <- subset(cars.a, cylinders!='' & cylinders!='-') # remove records wihtout cylinders data
cars.c <- subset(cars.b, displ!='' & displ!='-')         # remove records without displ data
#head(cars.c$cylinders)
cars.c$cylinders=as.numeric(as.character(cars.c$cylinders)) # convert cylinders to numeric
#head(cars.c$cylinders)
#head(cars.c$displ)
cars.c$displ=as.numeric(as.character(cars.c$displ))     # convert displ to numeric
#head(cars.c$displ)

summary(cars.c$fuelType)
summary(cars.c$fuelType1)
summary(cars.c$fuelType2)
summary(subset(cars.c[,c(31,32,68)], fuelType=='Gasoline or E85'))

library(plyr)
gastype = cars.c$fuelType
cars.c$gastype = mapvalues(gastype, from=c("CNG", "Gasoline or natural gas", "Gasoline or propane",
                          "Regular", "Midgrade", "Premium", "Diesel", "Electricity",
                          "Premium Gas or Electricity", "Regular Gas and Electricity",
                          "Gasoline or E85", "Premium or E85"), to = c("Regular", 
                          "Regular", "Regular", "Regular", "Regular", "Premium", "Diesel",
                          "Electricity", "Electricity", "Electricity", "E85", "E85"))

cars.d = ddf(cars.c)
time.glm = system.time(glm(UCity ~ UHighway + cylinders + displ + gastype + year, data = cars.c))
time.drGLM = system.time(drGLM(UCity ~ UHighway + cylinders + displ + gastype + year, data = cars.c))
time.drBLB = system.time({cars.d = ddf(cars.c, update=T)
cars.e <- divide(cars.d, by=rrDiv(17000), update=TRUE,postTrans=function(x) 
      x[,c("UCity", "UHighway", "cylinders", "displ", "gastype", "year")])

recombine(cars.e,
   apply=drBLB(
      statistic = function(x, weights)
         coef(glm(UCity ~ UHighway + cylinders + displ + gastype + year, 
            data=x, weights = weights)),
      metric = function(x)
         quantile(x, c(0.05, 0.95)),
      R = 100,
      n = nrow(cars.e)
   ),
   combine=combMean()
)

recombine(
    data = cars.e,
    apply = drGLM(UCity ~ UHighway+cylinders+displ+gastype+year),
    combine = combMeanCoef()
)})

time.glm
time.drGLM
time.drBLB
```

Guillermo Basulto-Elias
We got the same conclusion: dplyr is faster..
/```{r}
# Exercise 1 Andee Kaplan example
library(devtools)
#install_github('datadr','hafen')
library(datadr)
data(adult)

## turn adult into a ddf
(adultDdf<-ddf(adult, update=TRUE))
names(adultDdf)
library(ggplot2)
edTable<-summary(adultDdf)$education$freqTable
edTable$var

## Names
names(adultDdf)

## Plot
library(ggplot2)
edTable <- summary(adultDdf)$education$freqTable
edTable$var <- with(edTable, reorder(var, Freq, mean))
qplot(var, Freq, data = edTable, geom = "point") + coord_flip()


datadr_test <- function() {
    # make a preTransFn to group some education levels
    edGroups <- function(v) {
        v$edGroup <- as.character(v$education)
        v$edGroup[v$edGroup %in% c("1st-4th", "5th-6th")] <- "Some-elementary"
        v$edGroup[v$edGroup %in% c("7th-8th", "9th")] <- "Some-middle"
        v$edGroup[v$edGroup %in% c("10th", "11th", "12th")] <- "Some-HS"
        v
    }
    
    # divide by edGroup and filter out 'Preschool'
    byEdGroup <- divide(adultDdf, by = "edGroup", preTransFn = edGroups, filterFn = function(x) x$edGroup[1] != 
        "Preschool", update = TRUE)
    
    print(byEdGroup)
    
    # tabulate number of people in each education group
    edGroupTable <- recombine(byEdGroup, apply = nrow, combine = combRbind())
    print(edGroupTable)
    
    # order for plotting
    edGroupTable$edGroup <- with(edGroupTable, reorder(edGroup, val, mean))
    print(qplot(edGroup, val, data = edGroupTable, geom = "point") + coord_flip())
}
datadr_test()


### The same with plyr
library(dplyr)
dplyr_test <- function() {
    byEdGroup_dplyr <- adult %.% mutate(edGroup = ifelse(as.character(education) %in% 
        c("1st-4th", "5th-6th"), "Some-elementary", ifelse(as.character(education) %in% 
        c("7th-8th", "9th"), "Some-middle", ifelse(as.character(education) %in% 
        c("10th", "11th", "12th"), "Some-HS", as.character(education))))) %.% 
        filter(edGroup != "Preschool") %.% group_by(edGroup)
    print(byEdGroup_dplyr)
    
    edGroupTable_dplyr <- byEdGroup_dplyr %.% summarise(count = n())
    print(edGroupTable_dplyr)
    
    # order for plotting
    edGroupTable_dplyr$edGroup <- with(edGroupTable_dplyr, reorder(edGroup, 
        count, mean))
    print(qplot(edGroup, count, data = edGroupTable_dplyr, geom = "point") + 
        coord_flip())
}
dplyr_test()

library(profr)
system.time({
datadr_prof <- profr(datadr_test())
})

system.time({
dplyr_prof <- profr(dplyr_test())
})
```

For problem 3.
/```{r}
# map expression to emit max petal length for each k/v pair
minMap <- expression({
   for(curMapVal in map.values)
      collect("min", min(curMapVal$comb08))
      collect("make", curMapVal[which.min(curMapVal$comb08), "make"])
})


minReduce <- expression(
   pre={
      globalMin <- NULL
   },
   reduce={
      globalMin <- unlist(reduce.values)
   },
   post={
      collect(reduce.key, globalMin)
   }
)

minRes <- mrExec(cars.d,
   map = minMap,
   reduce = minReduce
)

minRes[["make"]]
minRes[["min"]]



```


Cleaning data before working with it in any case is important, but especially when working datadr format which is very specific. The ddo and ddf objects are the only objects that many of the datadr functions work well with. If your data isn't clean, it won't be workable. 

-->




